Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-01-11 PAPER.tex
Made josh edits until Section 3

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[figuresright]{rotating}
\usepackage{natbib}
\usepackage{ctable}
\usepackage{comment}
\citestyle{aa}

@@

Bootstrapping uses sampling with replacement to estimate a posterior distribution. For example, measurements (like power 
spectra) can be made from different samples of data. Each of these measurements is a different realization drawn from some underlying distribution, and realizations are correlated with each other to a degree set by the fraction of sampled points that are held in common 
between them. Hence, through the process of resampling and averaging along different axes, we can estimate error bars for 
our results which represent the underlying distribution of values that are allowed by our measurements.

@@

Foreground bias is perhaps one of the main factors limiting $21$ cm results. Foreground signals lie $\sim4$-$5$ orders of 
magnitude above the cosmological signal. Though there are many techniques proposed for removing foregrounds (see e.g. \citealt{vedantham_et_al2012}; \citealt{parsons_et_al2012a}; \citealt{parsons_et_al2012b}; \citealt{dillon_et_al2013a}; \citealt{wang_et_al2013}; \citealt{parsons_et_al2014}; \citealt{liu_et_al2014a}; \citealt{liu_et_al2014b}; \citealt{dillon_et_al2015}; \citealt{pober_et_al2016a}; \citealt{trott_et_al2016}), most 
experiments remain ultimately limited by residuals rather than noise.

For a particular 
baseline length, there is a maximum delay imposed on sources attached to the sky, which corresponds to the light-crossing time between two 
antennas in a baseline. For longer baselines, this value increases, producing what is known as ``the 
wedge"
\citep{datta_et_al2010, parsons_et_al2012b, vedantham_et_al2012, pober_et_al2013, thyagarajan_et_al2013, liu_et_al2014a, liu_et_al2014b, patil_et_al2017}. 
The wedge describes a region in $k$-space contaminated by smooth spectrum foregrounds, bounded by baseline length (which is proportional to $k_{\perp}$) and delay (which is 
proportional to $k_{\parallel}$). Properties of the wedge can be used to isolate and 
%remove 
avoid foregrounds, as done by A15, 
\citet{parsons_et_al2014}, and \citet{jacobs_et_al2015}.

Although smooth spectrum foregrounds preferentially show up at low delay, or low $k$ modes, their isolation within the wedge is not perfect.  In deep measurements, power spectrum measurements at $k_{\parallel}$ values beyond 
the length of a baseline are often still contaminated at a low level. This leakage, particularly at low $k$'s, can be attributed to 
convolution kernels associated with Fourier-transforming visibilities into delay-space. In other words, smooth-spectrum 
foregrounds appear as $\delta$-functions in delay-space, convolved by the Fourier transform of the source spectrum and the 
antenna response, both of which could smear out the foregrounds and cause leakage outside the wedge \citep[e.g.][]{ewall-wice_et_al2017}.

@@

Determining the source of detections at higher $k$'s is more difficult. In previous power spectrum results, these detections have been explained as instrumental systematics, particularly time-variable cross talk, RFI, and calibration errors (A15; \citealt{parsons_et_al2014}; \citealt{dillon_et_al2014}; \citealt{beardsley_et_al2016}; \citealt{patil_et_al2017}). In the next section (Section \ref{sec:JackknifeOverview}), we will present some tests that can help distinguish these excesses from that of EoR.

In addition to foreground bias, noise can also be responsible for positive power spectrum detections if thermal noise is 
multiplied by itself. Every $21$ cm visibility measurement contains thermal noise that is comprised of receiver and sky noise. 
We expect this noise to be independent between antennas and thus we can beat it down (increase sensitivity) by integrating 
longer, using more baselines, etc. However, the squaring of noise occurs when cross-multiplying visibilities, which is shown by 
the two copies of $\textbf{x}$ in Equation \eqref{eq:qhat}. If both copies of $\textbf{x}$ come from the same baseline and time, it 
can result in power spectrum measurements that are higher than those predicted by the thermal noise of the instrument. One 
way to avoid this type of noise bias is to avoid cross-multiplying data from the same baselines or days. This ensures that the 
two quantities that go into a measurement have separate noises that don't correlate with each other.

@@

The jackknife is a resampling technique in which a statistic (i.e. power spectrum) is computed in subsets of the data. These 
subsets are then compared to reveal systematics. In this section we define two main tests --- the null test and the traditional 
jackknife --- and explain how a power spectrum detection must pass each. We then highlight how these tests can be used to 
help distinguish between different sources of bias.

\begin{itemize}
\item \textbf{Null Test}: A null test is a type of jackknife test that removes the astronomical signal from data in order to 
investigate underlying systematics (e.g., see \citet{keating_et_al2016} for examples from intensity mapping that are closely related to our current application). For example, one can 
divide data into two subsets by separating odd and even Julian dates, or the first half of the observing season from the second. 
Subtracting the two removes signal that is common to both subsets, including foregrounds and EoR. The resulting power 
spectrum should be consistent with thermal noise estimates; if it is not, it suggests the presence of a systematic that differs 
from one of the data subsets to the other (i.e. doesn't get subtracted perfectly). 
\item \textbf{Traditional Jackknife}: In a more broad sense, it is important to perform many jackknife tests in order to instill 
confidence in a final result. A stable result must be steadfast throughout all jackknives no matter how the data is sliced. 
Jackknives can be taken along several different axes --- for example, one could start with a full dataset, and compute a new 
power spectrum every time as a day of data is removed, or a baseline is removed. This type of jackknife would reveal bias 
present only at certain LSTs (such as a foreground source), for example, or misbehaving baselines.
\end{itemize}

@@

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},width=\columnwidth]{plots/toy_bias2.png}
	\caption{Power spectrum estimates for $\textbf{x}_{1}$ and $\textbf{x}_{2}$, two jackknives of the toy model. They suggest 
the presence of a systematic in $\textbf{x}_{1}$ only, illustrating how jackknives can be used to tease out excesses. Clean 
measurements should remain consistent despite the jackknife taken.}
	\label{fig:toy_bias2}
\end{figure}

>>>>

\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[figuresright]{rotating}
\usepackage{natbib}
\usepackage{ctable}
\usepackage{comment}
\usepackage{titlesec}
\citestyle{aa}

@@

Bootstrapping uses sampling with replacement to estimate a posterior distribution. For example, measurements (like power 
spectra) can be made from different samples of data. Each of these measurements is a different realization drawn from some underlying distribution, and realizations are correlated with each other to a degree set by the fraction of sampled points that are held in common 
between them. Hence, through the process of resampling and averaging along different axes, we can estimate error bars for 
our results which represent the underlying distribution of values that are allowed by our measurements (\citealt{efron_tibshirani1994}; \citealt{andrae2010}).

@@

Foreground bias is perhaps one of the main factors limiting $21$ cm results. Foreground signals lie $\sim4$-$5$ orders of 
magnitude above the cosmological signal. Though there are many techniques proposed for removing foregrounds (see e.g. \citealt{vedantham_et_al2012}; \citealt{parsons_et_al2012a}; \citealt{parsons_et_al2012b}; \citealt{dillon_et_al2013a}; \citealt{wang_et_al2013}; \citealt{parsons_et_al2014}; \citealt{liu_et_al2014a}; \citealt{liu_et_al2014b}; \citealt{dillon_et_al2015}; \citealt{pober_et_al2016a}; \citealt{trott_et_al2016}), most 
experiments currently remain limited by residuals rather than noise, especially at low $k$.

For a particular 
baseline length, there is a maximum delay imposed on sources attached to the sky, which corresponds to the light-crossing time between two 
antennas in a baseline. For longer baselines, this value increases, producing what is known as ``the 
wedge"
\citep{datta_et_al2010, parsons_et_al2012b, vedantham_et_al2012, pober_et_al2013, thyagarajan_et_al2013, liu_et_al2014a, liu_et_al2014b, patil_et_al2017}. 
The wedge describes a region in $k$-space contaminated by smooth spectrum foregrounds, bounded by baseline length (which is proportional to $k_{\perp}$) and delay (which is 
proportional to $k_{\parallel}$). Properties of the wedge can be used to isolate and 
%remove 
avoid foregrounds, as done by A15, 
\citet{parsons_et_al2014}, \citet{dillon_et_al2014}, \citet{dillon_et_al2015}, \citet{jacobs_et_al2015}, \citet{beardsley_et_al2016}, and \citet{trott_et_al2016}.

Although smooth spectrum foregrounds preferentially show up at low delay, or low $k$ modes, their isolation within the wedge is not perfect.  In deep measurements, power spectrum measurements at $k_{\parallel}$ values beyond 
the length of a baseline are often still contaminated at a low level. This leakage, particularly at low $k$'s, can be attributed to 
convolution kernels associated with Fourier-transforming visibilities into delay-space. In other words, smooth-spectrum 
foregrounds appear as $\delta$-functions in delay-space, convolved by the Fourier transform of the source spectrum, the signal chain, and the 
antenna response, all of which could smear out the foregrounds and cause leakage outside the wedge \citep[e.g.][]{ewall-wice_et_al2017}.

@@

Determining the source of detections at higher $k$'s is more difficult. In previous power spectrum results, these detections have been explained as instrumental systematics, particularly time-variable cross talk, RFI, cable reflections, and calibration errors (A15; \citealt{parsons_et_al2014}; \citealt{dillon_et_al2014}; \citealt{beardsley_et_al2016}; \citealt{patil_et_al2017}). In the next section (Section \ref{sec:JackknifeOverview}), we will present some tests that can help distinguish these excesses from that of EoR.

In addition to foreground bias, noise can also be responsible for positive power spectrum detections if thermal noise is 
multiplied by itself. Every $21$ cm visibility measurement contains thermal noise that is comprised of receiver and sky noise. 
We expect this noise to be independent between antennas and thus we can beat it down (increase sensitivity) by integrating 
longer, using more baselines, etc. However, the squaring of noise occurs when cross-multiplying visibilities, which is shown by 
the two copies of $\textbf{x}$ in Equation \eqref{eq:qhat}. If both copies of $\textbf{x}$ come from the same baseline and time, it 
can result in power spectrum measurements that are higher than those predicted by the thermal noise of the instrument. One 
way to avoid this type of noise bias is to avoid cross-multiplying data from the same baselines or days. This ensures that the 
two quantities that go into a measurement have separate noises that don't correlate with each other. We also note that if the noise level is known, this type of bias can be subtracted off, though this procedure is argued to be dangerous by \citet{dillon_et_al2014}.

@@

The jackknife is a resampling technique in which a statistic (i.e. power spectrum) is computed in subsets of the data (\citealt{quenouille1949}; \citealt{tukey1958}). These 
subsets are then compared to reveal systematics. In this section we define two main tests --- the null test and the traditional 
jackknife --- and explain how a power spectrum detection must pass each. We then highlight how these tests can be used to 
help distinguish between different sources of bias.

\begin{itemize}
\item \textbf{Null Test}: A null test is a type of jackknife test that removes the astronomical signal from data in order to 
investigate underlying systematics (e.g., see \citet{keating_et_al2016} for examples from intensity mapping that are closely related to our current application). For example, one can 
divide data into two subsets by separating odd and even Julian dates, or the first half of the observing season from the second. 
Subtracting the two removes signal that is common to both subsets, including foregrounds and EoR. The resulting power 
spectrum should be consistent with thermal noise estimates; if it is not, it suggests the presence of a systematic that differs 
from one of the data subsets to the other (i.e. doesn't get subtracted perfectly). 
\item \textbf{Traditional Jackknife}: In a broader sense, it is important to perform many jackknife tests in order to instill 
confidence in a final result. A stable result must be steadfast throughout all jackknives no matter how the data is sliced. 
Jackknives can be taken along several different axes --- for example, one could start with a full dataset, and compute a new 
power spectrum every time as a day of data is removed, or a baseline is removed. This type of jackknife would reveal bias 
present only at certain LSTs (such as a foreground source), for example, or misbehaving baselines.
\end{itemize}

@@

\begin{figure}
	\centering
	\includegraphics[trim={0cm 0cm 0cm 0cm},width=\columnwidth]{plots/toy_bias2.png}
	\caption{Power spectrum estimates for $\textbf{x}_{1}$ and $\textbf{x}_{2}$, two jackknives of the toy model. They suggest 
the presence of a systematic in $\textbf{x}_{1}$ only (which is exactly what was put in), illustrating how jackknives can be used to tease out excesses. Clean 
measurements should remain consistent despite the jackknife taken.}
	\label{fig:toy_bias2}
\end{figure}
